---
title: "Final Project"
author: 'MSDS 6372: Chloe Barker and Jack Pehlke'
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
library(dplyr)
library(forcats)
library(knitr)
library(naniar)
library(tidyverse)
library(ggplot2)
library(GGally)
library(janitor)
library(olsrr)
library(ggthemes)
library(caret)
library(RANN)
library(showtext)
library(corrplot)
library(patchwork)
library(scales)
# Font for visuals
font_add(family = "Open Sans", regular = "OpenSans-SemiBold.ttf")
showtext_auto()
par(family = "Open Sans")
library(extrafont)
par(family = "Open Sans")
```

### Data Preparation
```{r}
heart_data <- read.csv("heart_2020_cleaned.csv", header = TRUE)
heart_data$HeartDisease <- factor(heart_data$HeartDisease, levels = c("No","Yes"), labels= c("No","Yes"))
heart_data$HeartDisease <- relevel(heart_data$HeartDisease, ref = "Yes") # "Yes" group the positive reference group
cat_vars <- c("Smoking", "AlcoholDrinking", "Stroke", "DiffWalking","Sex","AgeCategory", "Race", "Diabetic", "PhysicalActivity", "GenHealth","Asthma", "KidneyDisease", "SkinCancer")
num_vars <- c("BMI", "PhysicalHealth", "MentalHealth", "SleepTime","AgeCategory.num","GenHealth.num")
#Potential New Variables
genhealth_levels<- c("Poor", "Fair","Good", "Very good","Excellent")
age_levels<- c("18-24","25-29","30-34","35-39","40-44","45-49","50-54","55-59","60-64","65-69","70-74","75-79","80 or older")
heart_data <- heart_data %>%
  mutate(GenHealth.num = as.numeric(factor(GenHealth, levels = genhealth_levels, ordered = TRUE)),
    AgeCategory.num = as.numeric(factor(AgeCategory, levels = age_levels, ordered = TRUE)))
heart_data <- heart_data %>% mutate(across(all_of(cat_vars), as.factor))
heart_data <- heart_data %>% mutate(across(all_of(num_vars), as.numeric))
# Make new dataset
factor_cols <- sapply(heart_data,is.factor)# Identify factor columns
level_counts <- sapply(heart_data[,factor_cols, drop = FALSE], nlevels)# Count levels
ordered_factor_col_names <- names(level_counts)[order(level_counts)]# Order factor columns(by # of levels)
non_factor_col_names <- names(heart_data)[!factor_cols] # Get non-factor column names
new_col_order <- c(ordered_factor_col_names, non_factor_col_names) # Combine
data_sorted <- heart_data[,new_col_order] # Reorder the data frame columns
write.csv(data_sorted, "heart2020_data_sorted.csv",row.names = FALSE)
```

### Data Setup
```{r}
heart_data = read.csv("heart2020_data_sorted.csv",header=TRUE)
heart_data <- heart_data %>% mutate(across(all_of(cat_vars), as.factor))
heart_data <- heart_data %>% mutate(across(all_of(num_vars), as.numeric))
```

### Overall Dataset
```{r}
# Table of percent and number counts 
heart = heart_data %>% group_by(HeartDisease) %>% 
  summarise(n = n()) %>% 
  mutate(percent=round(n/sum(n),4)*100)
heart
```

### Data Split
```{r}
library(caret)
set.seed(4)
trainIndex <- createDataPartition(heart_data$HeartDisease, p = 0.8, list = FALSE) # to get balance of yes/no
trainHeart <- heart_data[trainIndex, ]
testHeart <- heart_data[-trainIndex, ]
write.csv(trainHeart, "trainHeart.csv",row.names = FALSE)
write.csv(testHeart, "testHeart.csv",row.names = FALSE)

trainHeart <- read.csv("trainHeart.csv")
testHeart <- read.csv("testHeart.csv")
trainHeart$HeartDisease <- factor(trainHeart$HeartDisease, levels = c("No","Yes"))
testHeart$HeartDisease <- factor(testHeart$HeartDisease, levels = c("No","Yes"))
table(trainHeart$HeartDisease)
table(testHeart$HeartDisease)
```

### EDA: Training Dataset
```{r}
# Table of percent and number counts 
heartTrain = trainHeart %>% group_by(HeartDisease) %>% 
  summarise(n = n()) %>% 
  mutate(percent=round(n/sum(n),4)*100)
heartTrain

# Percent of disease (yes/no)
ggplot(heartTrain,aes(x=HeartDisease,y=percent,colour=HeartDisease))+geom_bar(aes(fill=HeartDisease),show.legend=F,stat="identity")+ylab("Percentage") + scale_fill_manual(values = c("Yes" = "#CC0035", "No" = "#354CA1")) + theme_minimal(base_family = "Open Sans") + xlab("Heart Disease Status")

trainHeart$HeartDisease.num = as.numeric(ifelse(trainHeart$HeartDisease=="Yes",1,0))
numeric_data <- trainHeart %>% select(where(is.numeric))
numeric_data %>% filter(HeartDisease.num == "0") %>% summary()
numeric_data %>% filter(HeartDisease.num == "1") %>% summary()
```

### Correlation Plot and Table (Numeric Only)
```{r}
colors <- colorRampPalette(c("#354CA1", "white", "#CC0035"))(200)

# Correlation Table
numeric_data %>%
  cor() %>%
  round(2)

# Correlation Plot
numeric_data %>%
  cor() %>%
  corrplot(method = "color", type = "upper", col = colors, tl.col = "black", addCoef.col = "black",number.cex = 0.9)
```

### Scatterplot Matrix of Numeric Variables
```{r, eval=FALSE}
numeric_data$HeartDisease = trainHeart$HeartDisease
numeric_data$HeartDisease <- factor(numeric_data$HeartDisease,levels = c("No","Yes"))
numeric_data$HeartDisease <- relevel(numeric_data$HeartDisease, ref = "Yes")
numeric_datapair = numeric_data  %>% select(-HeartDisease.num)

smu_colors <- c("No" = "#354CA1", "Yes" = "#CC0035")
p <- ggpairs(numeric_datapair,mapping = aes(colour = HeartDisease),lower = list(continuous = wrap("points", alpha = 0.3, size = 0.1)))

p + scale_color_manual(values = smu_colors) + scale_fill_manual(values  = smu_colors)
```

### Box Plots of Numeric Predictors
```{r}
smu2 <- c("Yes"="#CC0035","No"="#B0B7C3")
a = trainHeart %>% ggplot(aes(x=BMI, y=HeartDisease, fill = HeartDisease)) + geom_boxplot(show.legend = FALSE)+ scale_fill_manual(values = smu2) + theme_minimal(base_family = "Open Sans")
b = trainHeart %>% ggplot(aes(x=SleepTime, y=HeartDisease, fill = HeartDisease)) + geom_boxplot(show.legend = FALSE)+ scale_fill_manual(values = smu2)+theme_minimal(base_family = "Open Sans")
c = trainHeart %>% ggplot(aes(x=PhysicalHealth, y=HeartDisease, fill = HeartDisease)) + geom_boxplot(show.legend = FALSE)+ scale_fill_manual(values = smu2)+theme_minimal(base_family = "Open Sans")
d = trainHeart %>% ggplot(aes(x=MentalHealth, y=HeartDisease, fill = HeartDisease)) + geom_boxplot(show.legend = FALSE)+ scale_fill_manual(values = smu2)+ theme_minimal(base_family = "Open Sans")
e = trainHeart %>% ggplot(aes(x=AgeCategory.num, y=HeartDisease, fill = HeartDisease)) + geom_boxplot(show.legend = FALSE)+ scale_fill_manual(values = smu2)+theme_minimal(base_family = "Open Sans")
f = trainHeart %>% ggplot(aes(x=GenHealth.num, y=HeartDisease, fill = HeartDisease)) + geom_boxplot(show.legend = FALSE)+ scale_fill_manual(values = smu2)+theme_minimal(base_family = "Open Sans")

# More visual differences
(e|a)/(c|f)

# Less visual differences
(d|b)
```

### Variables with more than two levels: EDA
```{r}
# g1: AgeCategory
g1<-trainHeart %>% 
  group_by(AgeCategory,HeartDisease) %>%
  summarise(cnt=n()) %>%
  mutate(perc=round(cnt/sum(cnt),4))%>%
  arrange(desc(perc))
g1

# g2: GenHealth
g2<-trainHeart %>% 
  group_by(GenHealth,HeartDisease) %>%
  summarise(cnt=n()) %>%
  mutate(perc=round(cnt/sum(cnt),4))%>%
  arrange(desc(perc))
g2

# g3: Diabetic
g3<-trainHeart %>% 
  group_by(Diabetic,HeartDisease) %>%
  summarise(cnt=n()) %>%
  mutate(perc=round(cnt/sum(cnt),4))%>%
  arrange(desc(perc))
g3

# g4: Race
g4 <- trainHeart %>%
  group_by(Race, HeartDisease) %>%
  summarise(cnt = n()) %>%
  mutate(perc = round(cnt / sum(cnt), 4)) %>%
  arrange(desc(perc))
g4
```

### Proportion plots grouped by HeartDisease
```{r}
g1g <- ggplot(g1[c(14:26),], aes(x = reorder(AgeCategory, -perc), y = perc, fill = AgeCategory)) +
  geom_bar(stat = "identity", show.legend = FALSE) + ylab("Proportion of Disease") +xlab("Age Category")+
  theme_minimal(base_family = "Open Sans") + theme(axis.text.x = element_text(angle = 45, hjust = 1))

g2g <- ggplot(g2[c(6:10),], aes(x = reorder(GenHealth, -perc), y = perc, fill = GenHealth)) +
  geom_bar(stat = "identity", show.legend = FALSE) + ylab("Proportion of Disease") + xlab("General Health") + theme_minimal(base_family = "Open Sans")

g3g <- ggplot(g3[c(5:8),], aes(x = reorder(Diabetic, -perc), y = perc, fill = Diabetic)) +
  geom_bar(stat = "identity", show.legend = FALSE) + ylab("Proportion of Disease") + xlab("Diabetic") +
  theme_minimal(base_family = "Open Sans") + theme(axis.text.x = element_text(angle = 45, hjust = 1))

g4g <- ggplot(g4[c(7:12),], aes(x = reorder(Race, -perc), y = perc, fill = Race)) +
  geom_bar(stat = "identity", show.legend = FALSE) +ylab("Proportion of Disease") + xlab("Race") +
  theme_minimal(base_family = "Open Sans") + theme(axis.text.x = element_text(angle = 45, hjust = 1))


(g1g | g2g) / (g3g | g4g)
```

### Binary Categorical Predictors EDA 
```{r}
# b1: Smoking
b1<-trainHeart %>% 
  group_by(Smoking,HeartDisease) %>%
  summarise(cnt=n()) %>%
  mutate(perc=round(cnt/sum(cnt),4))%>%
  arrange(desc(perc))
b1

# b2:AlcoholDrinking
b2<-trainHeart %>% 
  group_by(AlcoholDrinking,HeartDisease) %>%
  summarise(cnt=n()) %>%
  mutate(perc=round(cnt/sum(cnt),4))%>%
  arrange(desc(perc))
b2

# b3: Stroke
b3 <- trainHeart %>% 
  group_by(Stroke, HeartDisease) %>%
  summarise(cnt = n()) %>%
  mutate(perc = round(cnt / sum(cnt), 4)) %>%
  arrange(desc(perc))
b3

# b4: DiffWalking
b4 <- trainHeart %>% 
  group_by(DiffWalking, HeartDisease) %>%
  summarise(cnt = n()) %>%
  mutate(perc = round(cnt / sum(cnt), 4)) %>%
  arrange(desc(perc))
b4

# b5: Sex
b5 <- trainHeart %>% 
  group_by(Sex, HeartDisease) %>%
  summarise(cnt = n()) %>%
  mutate(perc = round(cnt / sum(cnt), 4)) %>%
  arrange(desc(perc))
b5

# b6: PhysicalActivity
b6 <- trainHeart %>% 
  group_by(PhysicalActivity, HeartDisease) %>%
  summarise(cnt = n()) %>%
  mutate(perc = round(cnt / sum(cnt), 4)) %>%
  arrange(desc(perc))
b6

# b7: Asthma
b7 <- trainHeart %>% 
  group_by(Asthma, HeartDisease) %>%
  summarise(cnt = n()) %>%
  mutate(perc = round(cnt / sum(cnt), 4)) %>%
  arrange(desc(perc))
b7

# b8: KidneyDisease
b8 <- trainHeart %>% 
  group_by(KidneyDisease, HeartDisease) %>%
  summarise(cnt = n()) %>%
  mutate(perc = round(cnt / sum(cnt), 4)) %>%
  arrange(desc(perc))
b8

# b9: SkinCancer
b9 <- trainHeart %>% 
  group_by(SkinCancer, HeartDisease) %>%
  summarise(cnt = n()) %>%
  mutate(perc = round(cnt / sum(cnt), 4)) %>%
  arrange(desc(perc))
b9
```

### Proportion plots grouped by HeartDisease
```{r}
b1b <- ggplot(b1[c(3:4),], aes(x = reorder(Smoking, -perc), y = perc, fill = Smoking)) +
  geom_bar(stat = "identity", show.legend = FALSE) + scale_fill_manual(values = smu2) +
  labs(x = "Smoking Status", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

b2b <- ggplot(b2[c(3:4),], aes(x = reorder(AlcoholDrinking, -perc), y = perc, fill = AlcoholDrinking)) + geom_bar(stat = "identity", show.legend = FALSE)+ scale_fill_manual(values = smu2) + labs(x = "Alcohol Drinking", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

b3b <- ggplot(b3[c(3:4),], aes(x = reorder(Stroke, -perc), y = perc, fill = Stroke)) + geom_bar(stat = "identity", show.legend = FALSE) + scale_fill_manual(values = smu2) + labs(x = "Stroke", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

b4b <- ggplot(b4[c(3:4),], aes(x = reorder(DiffWalking, -perc), y = perc, fill = DiffWalking)) + geom_bar(stat = "identity", show.legend = FALSE) + scale_fill_manual(values = smu2) + labs(x = "Difficulty Walking", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

b5b <- ggplot(b5[c(3:4),], aes(x = reorder(Sex, -perc), y = perc, fill = Sex)) + geom_bar(stat = "identity", show.legend = FALSE) + labs(x = "Sex", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

b6b <- ggplot(b6[c(3:4),], aes(x = reorder(PhysicalActivity, -perc), y = perc, fill = PhysicalActivity)) + geom_bar(stat = "identity", show.legend = FALSE) + scale_fill_manual(values = smu2) + labs(x = "Physical Activity", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

b7b <- ggplot(b7[c(3:4),], aes(x = reorder(Asthma, -perc), y = perc, fill = Asthma)) + geom_bar(stat = "identity", show.legend = FALSE) + scale_fill_manual(values = smu2) + labs(x = "Asthma", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

b8b <- ggplot(b8[c(3:4),], aes(x = reorder(KidneyDisease, -perc), y = perc, fill = KidneyDisease)) + geom_bar(stat = "identity", show.legend = FALSE) + scale_fill_manual(values = smu2) + labs(x = "Kidney Disease", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

b9b <- ggplot(b9[c(3:4),], aes(x = reorder(SkinCancer, -perc), y = perc, fill = SkinCancer)) +
  geom_bar(stat = "identity", show.legend = FALSE) + scale_fill_manual(values = smu2) +
  labs(x = "Skin Cancer", y = "Proportion with Heart Disease") + theme_minimal(base_family = "Open Sans")

(b1b | b7b |b9b) / (b3b | b4b| b8b)
(b6b|b2b)
b5b
```

### Categorical Variables: Heart Disease Proportions
```{r}
cat_vars <- c("Smoking", "AlcoholDrinking", "Stroke", "DiffWalking","Sex","AgeCategory", "Race", "Diabetic", "PhysicalActivity", "GenHealth","Asthma", "KidneyDisease", "SkinCancer")

library(purrr)
cat_var_stats <- map_dfr(cat_vars, function(v) {
  prop <- trainHeart %>% group_by(.data[[v]]) %>%
    summarize(p_heart = mean(HeartDisease == "Yes", na.rm = TRUE)) %>%
    summarize(max_diff = max(p_heart) - min(p_heart))
  tibble(var = v, max_diff = prop$max_diff)
})

cat_var_stats %>% arrange(desc(max_diff)) %>%
  slice_max(max_diff, n = 15) %>%
  mutate(var = fct_reorder(var, max_diff)) %>%
  ggplot(aes(max_diff, var)) +geom_col(fill = "#CC0035") + labs(x = "Max Difference in % Heart Disease (Yes)", y = "Variable", title = "Categorical Predictors with Strongest Variation") + theme_minimal(base_family = "Open Sans",base_size = 12)
```

### Scatterplot Matrix 
```{r}
head(trainHeart)
num.trainHeart = trainHeart[c(1,15:18)]
head(num.trainHeart)
num.trainHeart$HeartDisease <- factor(num.trainHeart$HeartDisease, levels = c("No","Yes"))

smu_colors <- c("No" = "#354CA1", "Yes" = "#CC0035")
p <- ggpairs(num.trainHeart,mapping = aes(color = HeartDisease, fill = HeartDisease),lower = list(continuous = wrap("points", alpha = 0.3, size = 0.1)),diag  = list(continuous = wrap("densityDiag", alpha = 0.4))) + theme(axis.text.x = element_text(size = 4), axis.text.y = element_text(size = 5))

p + scale_color_manual(values = smu_colors) + scale_fill_manual(values  = smu_colors)
```

## OBJECTIVE 1: Objective 1: Display the ability to perform EDA and build a logistic regression model for interpretation purposes using the training data. Perform your multiple logistic regression analysis and provide interpretation of the regression coefficients and confidence intervals for relevant predictor variables. 

### Logistic Regression Model
```{r}
trainHeart$HeartDisease <- factor(trainHeart$HeartDisease, levels = c("No","Yes"))

logistic <- glm(HeartDisease~PhysicalHealth + BMI + AgeCategory + GenHealth + Smoking + AlcoholDrinking + Stroke + DiffWalking + Sex + Race +Diabetic + PhysicalActivity + Asthma + KidneyDisease + SkinCancer + MentalHealth + SleepTime,data=trainHeart,family="binomial")
summary(logistic)
AIC(logistic) # AIC: 116151

simple <- glm(HeartDisease~PhysicalHealth + BMI + AgeCategory + GenHealth + Smoking + AlcoholDrinking + Stroke + DiffWalking + Sex + Race +Diabetic + Asthma + KidneyDisease + SkinCancer + MentalHealth + SleepTime,data=trainHeart,family="binomial")
#summary(simple)
AIC(simple)  # AIC: 115989.1

# Estimate for odds ratio
exp(cbind("Odds ratio" = coef(simple), confint.default(simple, level = 0.95)))

#1 Unit increase odds ratio: 
exp(0.0090482)

# VIF
library(car)
vif(simple)

# effect plots
library(sjPlot)
library(sjmisc)

a1=plot_model(simple,type="pred",terms=c("BMI[all]"))
a2=plot_model(simple,type="pred",terms=c("MentalHealth[all]"))
a3=plot_model(simple,type="pred",terms=c("PhysicalHealth[all]"))
a4=plot_model(simple,type="pred",terms=c("SleepTime[all]"))

(a1|a2)/(a3|a4)
```

## Objective 2:  Build 4-5 additional classification models to compare to your model in objective 1.  The goal of this objective is to build a model where prediction performance is prioritized.

### Scatter Plots for Interactions
```{r}
ggplot(trainHeart,aes(x=BMI,y=AgeCategory.num, color = HeartDisease))+geom_point(alpha = 0.5, size = 0.3)+ scale_color_manual(values = smu_colors)
ggplot(trainHeart,aes(x=PhysicalHealth,y=MentalHealth, color = HeartDisease))+geom_point(alpha = 0.5, size = 0.3)+ scale_color_manual(values = smu_colors)
ggplot(trainHeart,aes(x=MentalHealth,y=SleepTime, color = HeartDisease))+geom_point(alpha = 0.5, size = 0.3)+ scale_color_manual(values = smu_colors)
ggplot(trainHeart,aes(x=SleepTime,y=PhysicalHealth, color = HeartDisease))+geom_point(alpha = 0.5, size = 0.3)+ scale_color_manual(values = smu_colors)
```

### Feature Selection for Non-Additive Model
```{r}
fitControl<-trainControl(method="repeatedcv",number=5,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)
set.seed(1234)

# Feature Selection
logistic.fit1<-train(HeartDisease ~ GenHealth + AgeCategory + Sex + poly(BMI, 3) + poly(PhysicalHealth, 3) + poly(MentalHealth, 3) + poly(SleepTime, 3) + poly(BMI, 3):GenHealth + poly(PhysicalHealth, 3):GenHealth + poly(MentalHealth, 3):GenHealth + poly(SleepTime, 3):GenHealth + poly(BMI, 3):AgeCategory + poly(PhysicalHealth, 3):AgeCategory + poly(MentalHealth, 3):AgeCategory + poly(SleepTime, 3):AgeCategory + GenHealth:AgeCategory + GenHealth:Sex + AgeCategory:Sex + Stroke + KidneyDisease + Diabetic + DiffWalking + Smoking + AlcoholDrinking + Race + PhysicalActivity + Asthma + SkinCancer,data=trainHeart,
               method="glmnet",
               trControl=fitControl,
               metric="logLoss")
coef(logistic.fit1$finalModel,logistic.fit1$finalModel$lambdaOpt)

# Feature Selection After Dropping Features Above
logistic.fit2 <- train(HeartDisease ~ poly(BMI, 2) + poly(PhysicalHealth, 2) + poly(MentalHealth, 2) + poly(SleepTime, 3) + poly(BMI, 2):GenHealth + poly(PhysicalHealth, 2):GenHealth + poly(MentalHealth, 2):GenHealth + poly(SleepTime, 2):GenHealth + poly(BMI, 2):AgeCategory + poly(PhysicalHealth, 2):AgeCategory + poly(MentalHealth, 2):AgeCategory + poly(SleepTime, 3):AgeCategory + GenHealth:AgeCategory + GenHealth:Sex + AgeCategory:Sex + AgeCategory + GenHealth + Smoking + AlcoholDrinking + Stroke + DiffWalking + Sex + Race + Diabetic + PhysicalActivity + Asthma + KidneyDisease + SkinCancer,
  data = trainHeart,
  method = "glmnet",
  trControl = fitControl,
  metric = "logLoss")

logistic.fit2
coef(logistic.fit2$finalModel, logistic.fit2$finalModel$lambdaOpt)
```

### Complex Logistic Regression Model
```{r}
# Additive Model for Comparison
simple <- glm(HeartDisease~PhysicalHealth + BMI + AgeCategory + GenHealth + Smoking + AlcoholDrinking + Stroke + DiffWalking + Sex + Race +Diabetic + Asthma + KidneyDisease + SkinCancer + MentalHealth + SleepTime + PhysicalActivity ,data=trainHeart,family="binomial")

# Complex 1: 
complex1 <- glm(HeartDisease ~ GenHealth + AgeCategory + Sex + GenHealth:AgeCategory + GenHealth:BMI + AgeCategory:Sex + Stroke + KidneyDisease + Diabetic + DiffWalking + Smoking + AlcoholDrinking + Race + PhysicalActivity + Asthma + SkinCancer + BMI + PhysicalHealth + MentalHealth + SleepTime,data = trainHeart, family = "binomial")

# Complex 2:
complex2 <- glm(HeartDisease ~ poly(BMI, 2) + poly(PhysicalHealth, 2) + poly(MentalHealth, 2) + poly(SleepTime, 3) + poly(BMI, 2):GenHealth + poly(PhysicalHealth, 2):GenHealth + poly(MentalHealth, 2):GenHealth + poly(SleepTime, 2):GenHealth + poly(BMI, 2):AgeCategory + poly(PhysicalHealth, 2):AgeCategory + poly(MentalHealth, 2):AgeCategory + poly(SleepTime, 3):AgeCategory + GenHealth:AgeCategory + GenHealth:Sex + AgeCategory:Sex + AgeCategory + GenHealth + Smoking + AlcoholDrinking + Stroke + DiffWalking + Sex + Race + Diabetic + PhysicalActivity + Asthma + KidneyDisease + SkinCancer, data = trainHeart, family = "binomial")

AIC(simple)
AIC(complex1)
AIC(complex2)

# effect plots
library(sjPlot)
library(sjmisc)

a5=plot_model(complex2,type="pred",terms=c("BMI[all]"))
a6=plot_model(complex2,type="pred",terms=c("MentalHealth[all]"))
a7=plot_model(complex2,type="pred",terms=c("PhysicalHealth[all]"))
a8=plot_model(complex2,type="pred",terms=c("SleepTime[all]"))

(a5|a6)/(a7|a8)
```

### Sampling for KNN/QDA/LDA
```{r}
library(pROC)

#Sample 10%
set.seed(1234)
sample_pct <- 0.10
train_sample_size <- floor(nrow(trainHeart) * sample_pct)
test_sample_size <- floor(nrow(testHeart) * sample_pct)

train_indices <- sample(seq_len(nrow(trainHeart)), size = train_sample_size)
test_indices <- sample(seq_len(nrow(testHeart)), size = test_sample_size)

trainHeart_sample <- trainHeart[train_indices, ]
testHeart_sample <- testHeart[test_indices, ]

#Identify numeric columns present in BOTH sets
numeric_cols <- intersect(
  names(trainHeart_sample)[sapply(trainHeart_sample, is.numeric)],
  names(testHeart_sample)[sapply(testHeart_sample, is.numeric)]
)

#Extract only shared numeric columns from both sets
train_numeric <- trainHeart_sample[, numeric_cols]
test_numeric  <- testHeart_sample[, numeric_cols]

#Use preProcess to compute scaling from training data only
scale_params <- preProcess(train_numeric, method = c("center", "scale"))

#Apply scaling to both train and test
train_scaled <- predict(scale_params, train_numeric)
test_scaled  <- predict(scale_params, test_numeric)

#Add back the target variable
train_scaled$HeartDisease <- factor(trainHeart_sample$HeartDisease, levels = c("No", "Yes"))
test_scaled$HeartDisease  <- factor(testHeart_sample$HeartDisease, levels = c("No", "Yes"))

#Model Control
fitControl <- trainControl(method = "repeatedcv",
                           number = 5,
                           repeats = 1,
                           classProbs = TRUE,
                           summaryFunction = mnLogLoss)

```

### KNN/LDA/QDA Models
``` {r}
#KNN Model
set.seed(1234)
grid <- expand.grid(k = c(3,5,7)) 
knn.fit <- train(HeartDisease ~ PhysicalHealth + BMI + MentalHealth + SleepTime,
                 data = train_scaled,
                 method = "knn",
                 trControl = fitControl,
                 metric = "logLoss",
                 tuneGrid = grid)

#LDA Model
lda.fit <- train(HeartDisease ~ PhysicalHealth + BMI + MentalHealth + SleepTime,
                 data = train_scaled,
                 method = "lda",
                 trControl = fitControl,
                 metric = "logLoss")

#QDA Model
qda.fit <- train(HeartDisease ~ PhysicalHealth + BMI + MentalHealth + SleepTime,
                 data = train_scaled,
                 method = "qda",
                 trControl = fitControl,
                 metric = "logLoss")

```

### ROC/AUC 
```{r}
#ROC and AUC
actual <- test_scaled$HeartDisease
pred.knn <- predict(knn.fit, test_scaled, type = "prob")[, "Yes"]
pred.lda <- predict(lda.fit, test_scaled, type = "prob")[, "Yes"]
pred.qda <- predict(qda.fit, test_scaled, type = "prob")[, "Yes"]

roc.knn <- roc(actual, pred.knn, levels = c("No", "Yes"))
roc.lda <- roc(actual, pred.lda, levels = c("No", "Yes"))
roc.qda <- roc(actual, pred.qda, levels = c("No", "Yes"))

plot(roc.lda, col = "blue", lwd = 2, main = "ROC Curves: LDA, KNN, QDA")
plot(roc.knn, col = "red", lwd = 2, add = TRUE)
plot(roc.qda, col = "green", lwd = 2, add = TRUE)
legend("bottomright", legend = c("LDA", "KNN", "QDA"),
       col = c("blue", "red", "green"), lwd = 2)

cat("AUC LDA:", auc(roc.lda), "\n")
cat("AUC KNN:", auc(roc.knn), "\n")
cat("AUC QDA:", auc(roc.qda), "\n")

actual <- test_scaled$HeartDisease

#Visualize More Plots

# LDA
pred.lda <- predict(lda.fit, test_scaled, type = "prob")[, "Yes"]
roc.lda <- roc(actual, pred.lda, levels = c("No", "Yes"))
plot(roc.lda, col = "blue", lwd = 2, main = "ROC Curve: LDA")
cat("AUC LDA:", auc(roc.lda), "\n")

# KNN
pred.knn <- predict(knn.fit, test_scaled, type = "prob")[, "Yes"]
roc.knn <- roc(actual, pred.knn, levels = c("No", "Yes"))
plot(roc.knn, col = "red", lwd = 2, main = "ROC Curve: KNN")
cat("AUC KNN:", auc(roc.knn), "\n")

# QDA
pred.qda <- predict(qda.fit, test_scaled, type = "prob")[, "Yes"]
roc.qda <- roc(actual, pred.qda, levels = c("No", "Yes"))
plot(roc.qda, col = "green", lwd = 2, main = "ROC Curve: QDA")
cat("AUC QDA:", auc(roc.qda), "\n")
```

### Confusion Matrices 
```{r}
# Confusion Matrices
threshold <- 0.1 

library(caret)

# Logistic models 
simple_probs <- predict(simple, trainHeart, type = "response")
simple.pred <- factor(ifelse(simple_probs > threshold, "Yes", "No"), levels = c("No", "Yes"))

complex1_probs <- predict(complex1, trainHeart, type = "response")
complex1.pred <- factor(ifelse(complex1_probs > threshold, "Yes", "No"), levels = c("No", "Yes"))

complex2_probs <- predict(complex2, trainHeart, type = "response")
complex2.pred <- factor(ifelse(complex2_probs > threshold, "Yes", "No"), levels = c("No", "Yes"))

# LDA predictions
lda_probs <- predict(lda.fit, train_scaled, type = "prob")[, "Yes"]
lda.pred <- factor(ifelse(lda_probs > threshold, "Yes", "No"), levels = c("No", "Yes"))

# QDA predictions
qda_probs <- predict(qda.fit, train_scaled, type = "prob")[, "Yes"]
qda.pred <- factor(ifelse(qda_probs > threshold, "Yes", "No"), levels = c("No", "Yes"))

# KNN predictions 
knn_probs <- predict(knn.fit, train_scaled, type = "prob")[, "Yes"]
knn.pred <- factor(ifelse(knn_probs > threshold, "Yes", "No"), levels = c("No", "Yes"))

# Now create confusion matrices as before
cm.simple <- confusionMatrix(simple.pred, trainHeart$HeartDisease, positive = "Yes")
cm.complex1 <- confusionMatrix(complex1.pred, trainHeart$HeartDisease, positive = "Yes")
cm.complex2 <- confusionMatrix(complex2.pred, trainHeart$HeartDisease, positive = "Yes")
cm.lda <- confusionMatrix(lda.pred, train_scaled$HeartDisease, positive = "Yes")
cm.qda <- confusionMatrix(qda.pred, train_scaled$HeartDisease, positive = "Yes")
cm.knn <- confusionMatrix(knn.pred, train_scaled$HeartDisease, positive = "Yes")

# Print confusion matrices
cat("Confusion Matrix: Simple Logistic Regression\n")
print(cm.simple)
cat("\nConfusion Matrix: Complex Logistic Regression 1\n")
print(cm.complex1)
cat("\nConfusion Matrix: Complex Logistic Regression 2\n")
print(cm.complex2)
cat("\nConfusion Matrix: LDA\n")
print(cm.lda)
cat("\nConfusion Matrix: QDA\n")
print(cm.qda)
cat("\nConfusion Matrix: KNN\n")
print(cm.knn)

model_names <- c("Simple", "Complex1", "Complex2", "LDA", "QDA", "KNN")

accuracies <- c(
  cm.simple$overall["Accuracy"],
  cm.complex1$overall["Accuracy"],
  cm.complex2$overall["Accuracy"],
  cm.lda$overall["Accuracy"],
  cm.qda$overall["Accuracy"],
  cm.knn$overall["Accuracy"]
)

sensitivities <- c(
  cm.simple$byClass["Sensitivity"],
  cm.complex1$byClass["Sensitivity"],
  cm.complex2$byClass["Sensitivity"],
  cm.lda$byClass["Sensitivity"],
  cm.qda$byClass["Sensitivity"],
  cm.knn$byClass["Sensitivity"]
)

specificities <- c(
  cm.simple$byClass["Specificity"],
  cm.complex1$byClass["Specificity"],
  cm.complex2$byClass["Specificity"],
  cm.lda$byClass["Specificity"],
  cm.qda$byClass["Specificity"],
  cm.knn$byClass["Specificity"]
)

performance_table <- data.frame(
  Model = model_names,
  Accuracy = round(accuracies, 4),
  Sensitivity = round(sensitivities, 4),
  Specificity = round(specificities, 4)
)

print(performance_table)

cat("AUC LDA:", auc(roc.lda), "\n")
cat("AUC QDA:", auc(roc.qda), "\n")
cat("AUC KNN:", auc(roc.knn), "\n")

```

### Logistic Simple/Complex Model Metrics
```{R}
# Logistic Simple/Complex Model Metrics with 0.3 threshold
# Function to compute metrics for a given model and dataset
get_metrics <- function(model, data, positive_class = "Yes", threshold = 0.3) {
  # Predict probabilities
  probs <- predict(model, newdata = data, type = "response")
  
  # Convert to predicted class using custom cutoff
  preds <- ifelse(probs >= threshold, positive_class, "No")
  
  # Convert to factors with same levels as actuals
  actual <- factor(data$HeartDisease, levels = c("No", positive_class))
  preds  <- factor(preds, levels = c("No", positive_class))
  
  # Confusion matrix
  cm <- confusionMatrix(preds, actual, positive = positive_class)
  
  # Extract metrics
  sens <- cm$byClass["Sensitivity"]
  spec <- cm$byClass["Specificity"]
  ppv  <- cm$byClass["Pos Pred Value"]
  npv  <- cm$byClass["Neg Pred Value"]
  
  # Prevalence
  prevalence <- mean(actual == positive_class)
  
  # AUROC
  roc_obj <- roc(actual, probs, levels = c("No", positive_class))
  auc_val <- auc(roc_obj)
  
  return(data.frame(
    Sensitivity = sens,
    Specificity = spec,
    Prevalence = prevalence,
    PPV = ppv,
    NPV = npv,
    AUROC = auc_val
  ))
}

# Apply to each model
metrics_simple   <- get_metrics(simple,   trainHeart, threshold = 0.1)
metrics_complex1 <- get_metrics(complex1, trainHeart, threshold = 0.1)
metrics_complex2 <- get_metrics(complex2, trainHeart, threshold = 0.1)

# Combine results
results <- rbind(
  Simple   = metrics_simple,
  Complex1 = metrics_complex1,
  Complex2 = metrics_complex2
)

print(results)
```
