---
title: "Hospitalization Stays Project"
author: "Chloe Barker and Jack Korbitz"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(naniar)
library(tidyverse)
library(ggplot2)
library(GGally)
library(olsrr)
library(ggthemes)
library(caret)
library(RANN)
library(showtext)
library(corrplot)
library(patchwork)
```

# Import Data 
```{r}
hospital_data=read.csv("/Users/chloebarker/Downloads/Hospital-Data-Project/Final Project 1/HospitalDurations.csv",stringsAsFactors = TRUE)
# Factor conversion
hospital_data$Med.Sc.Aff = factor(hospital_data$Med.Sc.Aff, levels = c(1, 2), labels = c("Yes", "No"))
hospital_data$Region = factor(hospital_data$Region, levels = c(1,2,3,4), labels = c("NE", "NC", "S", "W"))
# Check structure - two factors
str(hospital_data)
# Missing data check - no missing values 
vis_miss(hospital_data)
```


## EDA: Overall Scatter Plots for Each Predictor
```{r}
long_data <- hospital_data %>% select(-ID, -Med.Sc.Aff, -Region) %>%
  pivot_longer(cols = -Lgth.of.Sty,  
    names_to = "Variable",
    values_to = "Value")

ggplot(long_data, aes(x = Value, y = log(Lgth.of.Sty))) +
  geom_point(alpha = 0.6, color = "#0596a2") + geom_smooth(linewidth=.7,se=FALSE, color = "#E2583E") +
  facet_wrap(~ Variable, scales = "free_x") +
  theme_minimal() +
  labs(title = "Length of Stay vs. Other Hospital Variables",
    x = "Value of Other Variables",
    y = "Length.of.Stay")
```

## EDA: Interactions
```{R}
# Infection Risk vs log(Length of Stay) by Region
ggplot(hospital_data, aes(x = Avg.Pat , y = log(Lgth.of.Sty), color = Region)) +geom_point() +geom_smooth(method="lm", se=FALSE)+labs(title = "Infection Risk vs. log(Length of Stay) by Region", x = "Infection Risk",y = "Log(Length of Stay)") + theme_minimal()

# Average Number of Patients vs log(Length of Stay) by Region
ggplot(hospital_data, aes(x = Avg.Pat , y = log(Lgth.of.Sty), color = Region)) +geom_point() +geom_smooth(method="lm", se=FALSE)+labs(title = "Average Number of Patients vs. log(Length of Stay) by Region", x = "Average Number of Patients Per Day",y = "Log(Length of Stay)") + theme_minimal()

# Average Number of Nurses vs log(Length of Stay) by Region
ggplot(hospital_data, aes(x = Avg.Nur , y = log(Lgth.of.Sty), color = Region)) +geom_point() +geom_smooth(method="lm", se=FALSE)+labs(title = "Average Number of Nurses vs. log(Length of Stay) by Region", x = "Average Number of Full-time Nurses",y = "Log(Length of Stay)") + theme_minimal()
```

## EDA: Correlation Matrix
```{r}
hospital_data$log.Lgth.of.Sty = log(hospital_data$Lgth.of.Sty)
numeric_hospital_data_log = hospital_data %>% select(where(is.numeric)) %>% select(-ID, -Lgth.of.Sty)
ggcorr(numeric_hospital_data_log,label = TRUE,label_size = 4,size = 4,hjust = 0.75) + labs(title = "Correlation Matrix - log(Length of Stay)") + theme_minimal()
```

## EDA: Categorical Variables
```{r}
# Tables of Categorical Variables
table(hospital_data$Region)
table(hospital_data$Med.Sc.Aff)

# Box Plots by Regions
ggplot(hospital_data, aes(x = Region, y = Lgth.of.Sty, fill = Region)) + geom_boxplot() + labs(title = "Length of Stay by Region", y = "Length of Hospital Stay") + scale_fill_manual(values = c("#b6e3f1", "#f69380","#0596a2","#d6b6c8"))+ theme_bw(base_size = 14)

# Box Plots by Med School Affiliation
ggplot(hospital_data, aes(x = Med.Sc.Aff, y = Lgth.of.Sty, fill = Med.Sc.Aff)) + geom_boxplot(color = "black") + scale_fill_manual(values = c("#b6e3f1", "#f69380")) + labs(title = "Length of Stay by Medical School Affiliation",x = "Medical School Affiliation",y = "Length of Hospital Stay") + theme_bw(base_size = 14)

# Histogram of Length of Stay faceted by Region
hospital_data %>% ggplot(aes(x = Lgth.of.Sty)) + geom_histogram(binwidth = 0.25, fill = "#00648f", color = "white") +labs(title = "Distribution of Length of Stay by Region",x = "Length of Stay",y = "Count") + facet_wrap(~Region, ncol = 1) +theme_bw(base_size = 12)

# Histogram of Length of Stay faceted by Med School Affiliation
hospital_data %>% ggplot(aes(x = Lgth.of.Sty)) + geom_histogram(binwidth = 0.25, fill = "#00648f", color = "white") +labs(title = "Distribution of Length of Stay by Medical School Affiliation",x = "Length of Stay",y = "Count") + facet_wrap(~Med.Sc.Aff, ncol = 1) +theme_bw(base_size = 12)
```

## Categorical Variable Exploration
```{r}
## Categorical Graph
cat_vars = c("Region", "Med.Sc.Aff")

# Compute Adjusted R²
cat_var_stats = map_df(cat_vars, function(v) {
  f = as.formula(paste("log.Lgth.of.Sty ~", v))
  mod = lm(f, data = hospital_data)
  tibble(var = v, adj_r2 = summary(mod)$adj.r.squared)
})

# Plot
cat_var_stats %>% arrange(desc(adj_r2)) %>% mutate(var = fct_reorder(var, adj_r2)) %>% ggplot(aes(adj_r2, var)) + geom_col(fill = "#e2583e") + labs(x = "Adjusted R²", y = "Categorical Predictor", title = "Adjusted R² of Categorical Predictors for log(Length of Stay)") + theme_minimal(base_size = 12) 

plot_box = function(df, var) {
  medians = df %>%
    group_by(.data[[var]]) %>%
    summarize(median_log_LOS = median(log.Lgth.of.Sty, na.rm = TRUE)) %>%
    arrange(median_log_LOS)
  df[[var]] = factor(df[[var]], levels = medians[[var]])
  
ggplot(df, aes(x = .data[[var]], y = log.Lgth.of.Sty)) + geom_boxplot(outlier.alpha = 0.2, width = 0.6, fill = "#00648f") + stat_summary(fun = median, geom = "point", shape = 21, size = 2, fill = "#f69380", color = "#00648f") + labs(title = paste("log(Length of Stay) vs.", var), y = "log(Length of Stay)") + theme_minimal(base_size = 12) + theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
}

plot_box(hospital_data, "Region")| plot_box(hospital_data, "Med.Sc.Aff")
```

## EDA: Numeric Variables
```{r}
names(hospital_data)
# Numeric predictor variables
num_vars = c("Age", "Inf.Risk", "Rad.Cul.Rate", "R.CX.ray.Rat", "N.Beds", "Avg.Pat", "Avg.Nur", "Pct.Ser.Fac")

# Calculate Adjusted R² for each numeric variable
num_var_stats = map_df(num_vars, function(v) {
  f = as.formula(paste("log.Lgth.of.Sty ~", v))
  mod = lm(f, data = hospital_data)
  tibble(var = v, adj_r2 = summary(mod)$adj.r.squared)
})

# Plot
num_var_stats %>% arrange(desc(adj_r2)) %>% mutate(var = fct_reorder(var, adj_r2)) %>% ggplot(aes(adj_r2, var)) + geom_col(fill = "#00648f") + labs(x = "Adjusted R²", y = "Predictor", title = "Adjusted R² of Numeric Predictors for log(Length of Stay)") + theme_minimal(base_size = 12)

## Specific Numeric Variable Graphs
plot_scatter = function(df, var) {
  ggplot(df, aes(x = .data[[var]], y = log.Lgth.of.Sty)) +
    geom_point(alpha = 0.5, color = "#00648f") +
    geom_smooth(se = FALSE, color = "#e2583e", linewidth = 1) +
    labs(title = paste("log(Length of Stay) vs", var), x = var, y = "log(Length of Stay)") +
    theme_minimal(base_size = 12)
}

(plot_scatter(hospital_data, "Age") | plot_scatter(hospital_data, "Avg.Nur")) /
(plot_scatter(hospital_data, "N.Beds") | plot_scatter(hospital_data, "Inf.Risk"))

(plot_scatter(hospital_data, "R.CX.ray.Rat") | plot_scatter(hospital_data, "Avg.Pat")) /
(plot_scatter(hospital_data, "Rad.Cul.Rate") | plot_scatter(hospital_data, "Pct.Ser.Fac"))
```

# EDA: Scatterplots Length of Stay vs Explanatory Variables
```{r}
library(tidyverse)
library(ggthemes)
library(plotly)

# Histogram of Length of Stay vs Age
ggplot(hospital_data, aes(x = Age, y = log(Lgth.of.Sty))) +
  geom_point(alpha = 0.6, color = "#00648f") +
  geom_smooth(method = "lm", color = "#E2583E", se = FALSE) +
  labs(title = "Age vs Log Length of Stay",
       x = "Age",
       y = "Log Length of Stay") +
  theme_minimal()

# Histogram of Length of Stay vs Infection Risk
ggplot(hospital_data, aes(x = Inf.Risk, y = log(Lgth.of.Sty))) +
  geom_point(alpha = 0.6, color = "#00648f") +
  geom_smooth(method = "lm" , color = "#E2583E", se = FALSE) +
  labs(title = "Infection Risk vs Log Length of Stay",
       x = "Inf.Risk",
       y = "Log Length of Stay") +
  theme_minimal()

# Histogram of Length of Stay vs Avg.Pat
ggplot(hospital_data, aes(x = Avg.Pat, y = log(Lgth.of.Sty))) +
  geom_point(alpha = 0.6, color = "#00648f") +
  geom_smooth(method = "lm", color = "#E2583E", se = FALSE) +
  labs(title = "Avg.Pat vs Log Length of Stay",
       x = "Avg.Pat",
       y = "Log Length of Stay") +
  theme_minimal()

# Histogram of Length of Stay vs Avg.Nur
ggplot(hospital_data, aes(x = Avg.Nur, y = log(Lgth.of.Sty))) +
  geom_point(alpha = 0.6, color = "#00648f") +
  geom_smooth(method = "lm", color = "#E2583E", se = FALSE) +
  labs(title = "Avg.Nur vs Log Length of Stay",
       x = "Avg.Nur",
       y = "Log Length of Stay") +
  theme_minimal()
```

## Interaction 3D Plot 
```{r}
# 3d ggplot with Y= Length of Stay, X= Age, Z= Avg.Nur
library(plotly)
plot_ly(hospital_data, 
        x = ~Age, 
        y = ~Lgth.of.Sty, 
        z = ~Inf.Risk, 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 6, 
                      color = ~Lgth.of.Sty, 
                      colorscale = "Plasma", 
                      showscale = TRUE, 
                      opacity = 0.9)) %>%
  layout(title = "Age, Length of Stay, and Avg.Nur",
         scene = list(xaxis = list(title = "Age"),
                      yaxis = list(title = "Log Length of Stay"),
                      zaxis = list(title = "Infection Risk")))
```

## Forward Stepwise Variable Selection
```{r warning= FALSE}
library(leaps)
library(patchwork)
library(ggplot2)
library(ggthemes)

Hospital_Duration <- read.csv('/Users/chloebarker/Downloads/Hospital-Data-Project/Final Project 1/HospitalDurations.csv', header = TRUE)
reg.fwd=regsubsets(log(Lgth.of.Sty) ~.
                  ,data=Hospital_Duration
                  ,method="forward"
                  ,nvmax=11)


summary(reg.fwd)$adjr2
summary(reg.fwd)$rss
summary(reg.fwd)$bic

# Extract summary stats
bics <- summary(reg.fwd)$bic
adjr2 <- summary(reg.fwd)$adjr2
rss <- summary(reg.fwd)$rss

# Create data frame for plotting
df <- data.frame(
  predictors = 1:length(bics),
  BIC = bics,
  AdjR2 = adjr2,
  RSS = rss
)

# BIC plot
p1 <- ggplot(df, aes(x = predictors, y = BIC)) +
  geom_line() +
  geom_point(aes(x = which.min(BIC), y = min(BIC)), color = "red", shape = 10, size = 3) +
  labs(x = "# of predictors", y = "BIC") +
  theme_stata()

# Adjusted R2 plot
p2 <- ggplot(df, aes(x = predictors, y = AdjR2)) +
  geom_line() +
  geom_point(aes(x = which.max(AdjR2), y = max(AdjR2)), color = "red", shape = 10, size = 3) +
  labs(x = "# of predictors", y = "Adjusted R-squared") +
  theme_stata()

# RSS plot
p3 <- ggplot(df, aes(x = predictors, y = RSS)) +
  geom_line() +
  geom_point(aes(x = which.min(RSS), y = min(RSS)), color = "red", shape = 10, size = 3) +
  labs(x = "# of predictors", y = "train RSS") +
  theme_stata()

# Combine all three plots in one row
p1 + p2 + p3

coef(reg.fwd,5) # Set to the BIC # of selected variables
```

## OBJ 1 MLR Model 1
```{r}
library(boot)
hospital_data = read.csv('/Users/chloebarker/Downloads/Hospital-Data-Project/Final Project 1/HospitalDurations.csv', header = TRUE)
hospital_data$Med.Sc.Aff = factor(hospital_data$Med.Sc.Aff, levels = c(1, 2), labels = c("Yes", "No"))
hospital_data$Region = factor(hospital_data$Region, levels = c(1,2,3,4), labels = c("NE", "NC", "S", "W"))
hospital_data$log.Lgth.of.Sty = log(hospital_data$Lgth.of.Sty)

set.seed(2025)
splitIndex = createDataPartition(hospital_data$log.Lgth.of.Sty, p = 0.8, list = FALSE)
train = hospital_data[splitIndex, ]
test = hospital_data[-splitIndex, ]

# Additivive model
model = lm(log.Lgth.of.Sty ~ Region + Avg.Pat + Age + Inf.Risk + Avg.Nur, data = train)
summary(model)
par(mfrow=c(2,2))
plot(model)
par(mfrow = c(1,1))

library(lmboot)

# Paired bootstrap
set.seed(1234)
boot.p <- paired.boot(log.Lgth.of.Sty ~ Region + Avg.Pat + Age + Inf.Risk + Avg.Nur, B = 1000, seed = 1234, data = train)

# 95% bootstrap percentile CI
ci_table = t(apply(boot.p$bootEstParam, 2, quantile, probs = c(0.025, 0.975)))
ci_table
```

# Variable Creation
```{r}
hospital_data2 = read.csv("/Users/chloebarker/Downloads/Hospital-Data-Project/Final Project 1/HospitalDurations.csv")
hospital_data2$Med.Sc.Aff = factor(hospital_data2$Med.Sc.Aff, levels = c(1, 2), labels = c("Yes", "No"))
hospital_data2$Region = factor(hospital_data2$Region, levels = 1:4, labels = c("NE", "NC", "S", "W"))
hospital_data2$log.Lgth.of.Sty = log(hospital_data2$Lgth.of.Sty)

hospital_data2 = hospital_data2 %>% mutate(
  Inf.Risk2 = Inf.Risk^2,
  Inf.Risk3 = Inf.Risk^3,
  Inf.Risk4 = Inf.Risk^4,
  N.Beds2 = N.Beds^2,
  N.Beds3 = N.Beds^3,
  N.Beds4 = N.Beds^4,
  Avg.Nur2 = Avg.Nur^2,
  Avg.Nur3 = Avg.Nur^3,
  Age2 = Age^2,
  Age3 = Age^3,
  Age4 = Age^4,
  Avg.Pat2 = Avg.Pat^2,
  Avg.Pat3 = Avg.Pat^3,
  Avg.Pat4 = Avg.Pat^4,
  Avg.Pat5 = Avg.Pat^5,
  R.CX.ray.Rat2 = R.CX.ray.Rat^2,
  R.CX.ray.Rat3 = R.CX.ray.Rat^3,
  R.CX.ray.Rat4 = R.CX.ray.Rat^4,
  Rad.Cul.Rate2 = R.Cul.Rat^2,
  Rad.Cul.Rate3 = R.Cul.Rat^3,
  Rad.Cul.Rate4 = R.Cul.Rat^4,
  Rad.Cul.Rate5 = R.Cul.Rat^5,
  Pct.Ser.Fac2 = Pct.Ser.Fac^2,
  Pct.Ser.Fac3 = Pct.Ser.Fac^3,
  Pct.Ser.Fac4 = Pct.Ser.Fac^4)

hospital_data2 <- hospital_data2 %>% select(-ID, -Lgth.of.Sty)
```

## Lasso Regression for Model 2
```{r}
x=model.matrix(log.Lgth.of.Sty~.,hospital_data2)[,-1]
y=hospital_data2$log.Lgth.of.Sty

library(glmnet)
set.seed(1234)

cv.out=cv.glmnet(x,y,alpha=1)
plot(cv.out)
bestlambda<-cv.out$lambda.1se
coef(cv.out,s=bestlambda)
```

## OBJ 2: Recursive Feature Elimination (RFE) for Model 2
```{r}
# Train/test split
set.seed(2025)
splitIndex = createDataPartition(hospital_data2$log.Lgth.of.Sty, p = 0.8, list = FALSE)
train2 = hospital_data2[splitIndex,]
test2 = hospital_data2[-splitIndex,]

# Model 2: Random Forest
# Define the control using a random forest selection function
ctrl <- rfeControl(functions = rfFuncs, method = "boot")

# Run RFE
rfe_result <- rfe(
  x = train2[, setdiff(names(train2), "log.Lgth.of.Sty")],
  y = train2$log.Lgth.of.Sty,
  sizes = c(5, 10, 15, 20),  # Number of variables to test
  rfeControl = ctrl
)

# View top predictors
predictors(rfe_result)
```

## OBJ 2 MLR Model 2
```{r}
# Complex MLR
model_complex= lm(log.Lgth.of.Sty ~ 
                        Inf.Risk * Region +
                        Avg.Pat * Region +
                        Avg.Nur * Region +
                        Age2 +
                        Avg.Pat4 + Avg.Pat3 +
                        Rad.Cul.Rate5 + 
                        Pct.Ser.Fac + Pct.Ser.Fac3 +  
                        R.CX.ray.Rat + R.CX.ray.Rat2 + R.CX.ray.Rat3 + R.CX.ray.Rat4 , data = train2)

summary(model_complex)

par(mfrow = c(2, 2))
plot(model_complex, main = "Complex Model Residuals")
par(mfrow = c(1, 1))

# Residual plot
plot(model_complex$residuals, main = "Residuals Plot", ylab = "Residuals", xlab = "Index", col = "#354CA1")
```

## OBJ 2 NON-parametric Model 3: KNN Model
```{r}
hospital_data1 = hospital_data2
set.seed(2025)
hospital_data1$log.Lgth.of.Sty = log(hospital_data1$Lgth.of.Sty)
splitIndex = createDataPartition(hospital_data1$log.Lgth.of.Sty, p = 0.8, list = FALSE)
train1 = hospital_data1[splitIndex,]
test1 = hospital_data1[-splitIndex,]


fitControl<-trainControl(method="repeatedcv",number=10,repeats=10) 
knn.fit<-train(log.Lgth.of.Sty ~ Region * Avg.Pat + Age + Inf.Risk + Avg.Nur + R.CX.ray.Rat,
                  data=train1,
                  method="knn",
                  trControl=fitControl)
print(knn.fit)
plot(knn.fit)
knn.fit

knn.fit2<-train(log.Lgth.of.Sty ~ Region * Avg.Pat + Age + Inf.Risk + Avg.Nur + R.CX.ray.Rat,
               data=train1,
               method="knn",
               trControl=fitControl,
               tuneGrid=expand.grid(k=c(1:10,20,30))
)

knn.fit2
plot(knn.fit2)
knn.fit2$results

```

# Overall Comparison of Metrics 
```{r}
set.seed(123)
splitIndex = createDataPartition(hospital_data2$log.Lgth.of.Sty, p = 0.8, list = FALSE)
train2 = hospital_data2[splitIndex, ]
test2 = hospital_data2[-splitIndex, ]

base_model = lm(log.Lgth.of.Sty ~ Region + Avg.Pat + Age + Inf.Risk + Avg.Nur, data = train2)
summary(base_model)

complex_model = lm(log.Lgth.of.Sty ~ 
                        Inf.Risk * Region +
                        Avg.Pat * Region +
                        Avg.Nur * Region +
                        Age + Age2 +
                        Avg.Pat4 + Avg.Pat3 + Avg.Pat5 +
                        Rad.Cul.Rate5 + 
                        Pct.Ser.Fac + Pct.Ser.Fac3 +  
                        R.CX.ray.Rat + R.CX.ray.Rat2 + R.CX.ray.Rat3 + R.CX.ray.Rat4 , data = train2)
summary(complex_model)


# Predictions
pred_base = predict(base_model, newdata = test2)
pred_complex = predict(complex_model, newdata = test2)
pred_knn = predict(knn.fit2, newdata = test1)

# MSE Comparison
mse_base = mean((test2$log.Lgth.of.Sty - pred_base)^2)
mse_complex = mean((test2$log.Lgth.of.Sty - pred_complex)^2)
mse_knn = mean((test1$log.Lgth.of.Sty - pred_knn)^2)

# RMSE
rmse_base = sqrt(mse_base)
rmse_complex = sqrt(mse_complex)
rmse_knn = sqrt(mse_knn)

# Adjusted R² Comparison
adj_r2_base = summary(base_model)$adj.r.squared
adj_r2_complex = summary(complex_model)$adj.r.squared
adj_r2_knn = cor(test1$log.Lgth.of.Sty, pred_knn)^2  # R² for k-NN

# MAE
mae_base = mean(abs(test2$log.Lgth.of.Sty - pred_base))
mae_complex = mean(abs(test2$log.Lgth.of.Sty - pred_complex))
mae_knn = mean(abs(test1$log.Lgth.of.Sty - pred_knn))

# AIC
aic_base = AIC(base_model)
aic_complex = AIC(complex_model)
aic_knn = ""  # AIC for k-NN is not directly available, using model fit

# BIC
bic_base = BIC(base_model)
bic_complex = BIC(complex_model)
bic_knn = ""  # BIC for k-NN is not directly available, using model fit

# Table of Results
model_compare = tibble(
  Model = c("Base MLR", "Complex MLR", "k-NN"),
  MSE = c(mse_base, mse_complex, mse_knn),
  RMSE = c(rmse_base, rmse_complex, rmse_knn),
  Adj_R2 = c(adj_r2_base, adj_r2_complex, adj_r2_knn),
  MAE = c(mae_base, mae_complex, mae_knn),
  AIC = c(aic_base, aic_complex, aic_knn),
  BIC = c(bic_base, bic_complex, bic_knn)
)

print(model_compare)
```

## APPENDIX
## Random Forest Model
```{r}
# Train/test split
set.seed(2025)
splitIndex = createDataPartition(hospital_data2$log.Lgth.of.Sty, p = 0.8, list = FALSE)
train2 = hospital_data2[splitIndex,]
test2 = hospital_data2[-splitIndex,]

# Model 2: Random Forest
# Define the control using a random forest selection function
ctrl <- rfeControl(functions = rfFuncs, method = "boot")

# Run RFE
rfe_result <- rfe(
  x = train2[, setdiff(names(train2), "log.Lgth.of.Sty")],
  y = train2$log.Lgth.of.Sty,
  sizes = c(5, 10, 15, 20),  # Number of variables to test
  rfeControl = ctrl
)

# View top predictors
predictors(rfe_result)

# Train model with selected variables and Region interaction
selected_vars <- predictors(rfe_result)
formula <- as.formula(
  paste("log.Lgth.of.Sty ~ (", paste(selected_vars, collapse = " + "), ") * Region")
)

model2 <- train(formula,
  data = train2,
  method = "rf",
  tuneLength = 5,
  trControl = trainControl(method = "boot")
)

print(model2)


```
